% -*- coding: utf-8 -*-
% This is part of the book TeX for the Impatient.
% Copyright (C) 2003 Paul W. Abrahams, Kathryn A. Hargreaves, Karl Berry.
% See file fdl.tex for copying conditions.

\input macros
%\chapter{Using this book}
\chapter{使用本书}

\chapterdef{usebook}

%This book is a do-it-yourself guide and handbook for \TeX.
%Here in this section we tell you how to use the book to maximum advantage.

本书是一本介绍 \TeX{} 的自学手册。
在本章中, 我们将指导你如何更好地使用本书。

% We recommend that you first either read or skim in sequence Sections
% \chapternum{usebook} through \chapternum{examples},
% which tell you what you need to know in order to get started using \TeX.
% If you've already had experience using \TeX, it will still be helpful
% to know what kinds of information are in these sections of the book.
% Sections~\chapternum{concepts}--\chapternum{tips}, which
% occupy most of
% the rest of the book, are designed to be accessed randomly.
% Nevertheless, if you're the kind of person who likes to read reference manuals,
% you'll find that it \emph{is} possible to proceed sequentially if you're
% willing to take a lot of detours at first.

我们建议你先阅读或浏覧 \chapternum{usebook} 到 \chapternum{examples}
章，
这些地方会告诉你如何上手 \TeX。
如果你已经使用过 \TeX， 这些地方仍会值得一看，
至少你可以知道这些章节介绍了些什么内容。
%本书的重头戏是\chapternum{concepts}--\chapternum{tips}章,
%这些章节可以任意选择地进行阅读.
其馀部分，主要是\chapternum{concepts}--\chapternum{tips}章，
可以有选择地翻阅。
当然， 如果你是一个很喜爱阅读参考手册的人，
你会发现从头到尾地把这本书阅读一遍\emph{是}完全可行的，
当然你会为此在开头学习时绕些弯路。

% In \chapterref{usingtex}, ``Using \TeX'',
% we explain how to produce a \TeX\ document from a
% \TeX{} input file.
% We also describe the conventions for preparing that input file,
% explain a little about how \TeX{} works, and tell you about additional
% resources that are available.
% Reading this section will help you understand
% the examples in the next section.

在 \chapterref{usingtex}， ``使用 \TeX''，
我们解释如何把一个 \TeX{} 输入文件转为一个 \TeX{} 文档。
%我们还讲述了准备这个输入文件的惯例，
我们还讲述了准备输入文件的一些习惯，
简要地解释了 \TeX{} 是如何工作的， 告诉你可以获取的其它资源信息。
阅读了这个章节， 会有助于你理解下一个章节的例子。

% \chapterref{examples}, ``Examples'',
% contains a sequence of
% examples that illustrate the use of \TeX.
% Each example consists of a page of output
% together with the input that we used to create it.
% These examples will orient you and help you
% locate the more detailed material that you'll need as you go.
% By seeing which commands are used in the input, you'll know
% where to look for more detailed information on how to achieve
% the effects shown in the output.
% The
% examples can also serve as models for simple documents, although we must
% caution you that
% because we've tried to pack a variety of \TeX\ commands into a small
% number of pages,
% the examples are not necessarily illustrations of good or complete
% document~design.

\chapterref{examples}, ``例子'',
包括了一系列的例子来描述 \TeX{} 的功用.
每个例子包括了一页的输出内容, 也包括了输出该页内容所输入的源代码.
这些例子会指导你且帮助你找到你所需要的更详细的资料.
通过观察输入使用了什么命令,
你会知道在哪里能够找到输出效果所使用方法的更详细的信息.
这些例子也可以作为简单的文档的模版, 当然我们必需提醒你,
因为我们尝试把很多的 \TeX{} 命令放在这几页纸中,
这些例子不可能描述好的或者完整的文档设计.

% As you read the explanation of a command, you may encounter
% some unfamiliar technical terms.
% In \chapterref{concepts}, ``Concepts'',
% we define and explain these terms.
% We also discuss other topics that aren't covered
% elsewhere in the book.
% The inside back cover of the book contains a list of all the
% concepts and the pages where they are described.
% We suggest that you make a copy of this list and keep it nearby
% so that you'll be able to identify and look up an unfamiliar
% concept immediately.

当你看这些命令的解释时, 你会遇到很多不熟悉的技术术语.
在 \chapterref{concepts}, ``概念''中, 我们定义并解释了这些术语.
我们同时还讨论了本书其它地方没有涉及到的话题.
本书书背内页列出了所有的概念名称以及它们所出现的位置.
我们建议你把书背内页复印下来, 放在旁边,
这样你就能在遇到不熟悉的概念时马上查找到它们.

% \TeX's commands are its primary vocabulary,
% and the largest part of this book is
% devoted to explaining them.  In Sections~\chapternum{paras}
% through~\chapternum{general} we describe the commands.
% You'll find general information about the command descriptions
% on \xrefpg{cmddesc}.
% The command descriptions are arranged
% functionally, rather like a thesaurus, so if you know what you want to
% do but you don't know which command does it for you, you can use the
% table of contents to guide you to the right group of commands.
% Commands that we think are both particularly useful and easy to understand
% are indicated with a pointing~hand~(\hand).

\TeX{} 的命令是它的主要词汇, 本书最主要的部分就是用来解释这些命令.
从第~\chapternum{paras}~章到第~\chapternum{general}~章, 我们解释了这些命令.
你可以在 \xrefpg{cmddesc} 找到命令描述的描述惯例.
这些命令根据功能分门别类, 而不是按字典顺序,
所以如果你知道你想做什么, 但不知那该用什么命令,
你可以使用目录来引导你找到正确的命令组.
我们给我们认为常用且容易理解的命令标上了一个手形符号~(\hand).


\chapterref{capsule}, ``Capsule summary of commands'', is a
specialized index that complements the more complete descriptions
in Sections~\chapternum{paras}--\chapternum{general}.
It lists \TeX's commands
alphabetically, with a brief explanation of each command
and a reference to the page
where it is described more completely.  The capsule summary
will help you when you just want a quick reminder of what a command
does.

\TeX{} is a complex program that occasionally works its will in
mysterious ways.
In \chapterref{tips}, ``Tips and techniques'',
we provide advice on solving a variety of specific
problems that you may encounter from time to time.
And if you're stumped by
\TeX's error messages, you'll find succor in \chapterref{errors},
``Making sense of error messages''.

The gray tabs on the side of the book will help you locate parts of the
book quickly.  They divide the book into the following major parts:
\olist
\li general explanations and examples
\li concepts
\li descriptions of commands (five shorter tabs)
\li advice, error messages, and the |eplain.tex| macros
\li capsule summary of commands
\li index
\endolist

In many places we have provided page references to
^{\texbook} (see \xrefpg{resources} for a citation).
These references apply to the seventeenth edition of \texbook.
For other editions, some references may be off by a page or two.


\section Syntactic conventions

In any book about preparing input for a computer,
it's necessary to indicate clearly the literal characters that should be typed
and to distinguish those characters from the explanatory text.
We use the Computer Modern typewriter font for {\tt literal input
like this}, and also for the names of \TeX{} commands.
When there's any possibility of confusion, we enclose \TeX{}
input in single quotation marks, `{\tt like this}'.
However, we occasionally use parentheses when we're indicating single
characters such as (|`|) (you can see why).

For the sake of your eyes we usually just put spaces
where you should put spaces. In some places where
we need to emphasize the space, however,
we use a `\visiblespace' character
{\catcode `\ =\other\pix^^| |}%
to indicate it.
Naturally enough, this character is called a \emph{visible space}.
\pix^^{spaces//visible}


\section Descriptions of the commands

\xrdef{cmddesc}
Sections~\chapternum{paras}--\chapternum{general} contain
a description of what nearly every \TeX{} command does.  ^^{commands}
Both the primitive commands ^^{primitive//command}
and those of ^{\plainTeX} are covered.
The primitive commands are those built into the \TeX{} computer program, while
the \plainTeX{} commands are defined in a standard file of
auxiliary definitions (see \xref\plainTeX).
The only commands we've omitted are those that are used purely locally
in the definition of \plainTeX\ (\knuth{Appendix~B}).
The commands are organized as follows:
\ulist\compact
\li ``Commands for composing paragraphs'', \chapterref{paras},
deal with characters, words, lines, and entire paragraphs.
\li ``Commands for composing pages'', \chapterref{pages},
deal with pages, their components, and the output routine.
\li ``Commands for horizontal and vertical modes'', \chapterref{hvmodes},
have corresponding or identical
forms for both horizontal modes (paragraphs and hboxes) and vertical
modes (pages and vboxes).
These commands provide boxes, spaces, rules, leaders,
and alignments.
\li ``Commands for composing math formulas'', \chapterref{math},
provide capabilities for constructing math formulas.
\li ``Commands for general operations'', \chapterref{general},
provide
\TeX's programming features and
everything else that doesn't fit into any of the other sections.
\endulist
You should think of these categories as being suggestive rather than
rigorous, because the commands don't really fit neatly into these
(or any other) categories.

Within each section, the descriptions of the commands are organized
by function.  When several commands are closely related, they are described as
a group; otherwise, each command has its own explanation.
The description of each command
includes one or more examples and the output
produced by each example when examples are appropriate (for
some commands they aren't).
When you are looking at a subsection containing functionally related
commands, be sure to check the end of a subsection for a ``see also''
item that refers you to related commands that are described elsewhere.

Some commands are closely related to certain concepts.
For instance, the |\halign| and |\valign| commands are related to
``alignment'', the |\def|
command is related to ``macro'',
and the |\hbox| and |\vbox| commands are related to ``box''.
In these cases we've usually given a bare-bones des\-crip\-tion of the
commands themselves and explained  the underlying ideas
in the concept.

The examples associated with the commands have been typeset with
^|\parindent|, the paragraph indentation, set to zero so that
paragraphs are normally unindented.
This convention makes the examples easier to read.
In those examples where the paragraph indentation is essential,
we've set it explicitly to a nonzero value.

The pointing hand in front of a command or a group of commands indicates
that we judged this command or group of commands to be particularly useful
and easy to understand.

Many commands expect ^{arguments} of one kind or another
(\xref{arg1}).  The arguments of
a command give \TeX{} additional information that it needs in order to
carry out the command.  Each argument is indicated by an italicized
term in angle brackets that indicates what kind of argument it~is:

\display{%
\halign{\<#>\quad&#\hfil\cr
argument&a single token or some text enclosed in braces\cr
charcode&a character code, i.e., an integer between $0$ and $255$\cr
dimen&a dimension, i.e., a length\cr
glue&glue (with optional stretch and shrink)\cr
number&an optionally signed integer (whole number)\cr
register&a register number between $0$ and $255$\cr
}}
^^{\<dimen>}
^^{\<argument>}
^^{\<charcode>}
^^{\<glue>}
^^{\<number>}
^^{\<register>}

% \noindent
% All of these terms are explained in more detail in \chapterref{concepts}.
% In addition, we sometimes use terms such as \<token list> that are either
% self-explanatory or explained in the description of the command.
% Some commands have special formats that require either braces or
% particular words.
% These are set in the same bold font that we use
% for the command headings.

\noindent
所有这些术语会在 \chapternum{concepts} 中进行详述.
除了这些以外, 我们会使用一些不言自明的或者在后文进行讲述的述语, 比如 \<token list>.
有些命令有自己独特的格式, 比如需要括孤或者特定的字词.
它们会使用和命令标题相同的黑体标出.

% Some commands are parameters (\xref{introparms}) or table entries.
% ^^{parameters//as commands}
% This is indicated in the command's listing.
% You can either use a parameter as an argument or assign a value to it.
% The same holds for table entries.
% We use the term ``parameter'' to refer to entities such as |\pageno|
% that are actually registers but behave just like parameters.
% ^^{registers//parameters as}

有些命令是参项 (\xref{introparms}) 或表项.
^^{parameters//as commands}
这会在命令的列表中指示出来.
你可以作为一个命令的参数使用这个参项, 也可以给它赋值.
表项也是如此.
我们使用``参项''这个术语来指代这些条目, 比如本是寄存器但可当参项使用的 |\pageno|.
^^{registers//parameters as}


\endchapter
\byebye
