~% -*- coding: utf-8 -*-
~% This is part of the book TeX for the Impatient.
~% Copyright (C) 2003 Paul W. Abrahams, Kathryn A. Hargreaves, Karl Berry.
~% See file fdl.tex for copying conditions.
% TeX ignores anything on a line after a %
% The next two lines define fonts for the title
\font\xmplbx = cmbx10 scaled \magstephalf
\font\xmplbxti = cmbxti10 scaled \magstephalf
% Now here's the title.
\leftline{\xmplbx Example !xmpnum:\quad\xmplbxti Entering simple text}
\vglue .5\baselineskip % skip an extra half line
~\count255 = \pageno
~\xdef\examplepage{\number\count255}
~\markinfo{Example 1: Entering simple text}
~\ifrewritetocfile
~\write\tocfile{\string\tocsectionentry{Entering simple text}{}{\examplepage}}%
~\fi
~^^{footnotes} ^^{comments} ^^{punctuation} ^^{quotation marks}
~^^{space characters} ^^{dashes} ^^{paragraphs//ending}
~\edef\examplepageno{\number\count255}%
%It's easy to prepare ordinary text for \TeX\ since
%\TeX\ usually doesn't care about how you break up your input into
%lines. It treats the end of a line of text like a space.%
%\footnote \dag{\TeX\ treats a tab like a space too, as we point
%out in this {\it footnote}.}   If you don't want a space there,
%put a per%
%cent sign (the comment character) at the end of the line.
%   \TeX\ ignores spaces at the start of a line, and treats more
%than        one      space as equivalent to a single space,
%even after a period.      You indicate a new paragraph by
%skipping a line (or more than one line).

由于 \TeX\ 并不关心原始的输入文本中的换行，因此，我们无须关心
应该给其提供何种格式化的文本。在 \TeX\ 看来，输入文本中的换行符
和空格等价。%
\footnote \dag{\TeX\ 将 Tab 也当作空格处理, 就如你在这个
{\it footnote}所见到的一样。}   如果你不想在行尾处看到空格，
可以在行尾处写一个百分号（注释符号）。\TeX\ 忽略行首的空格，
对多于一个的空格当做一个单独的空格看待 --- 哪怕这些空格位于
句号后面。你可以用一个或者多个空白行表示一个新的段落的开始。


%When \TeX\ sees a period followed by a space (or the end of the
%line, which is equivalent), it ordinarily assumes you've ended a
%sentence and inserts a little extra space after the period.  It
%treats question marks and exclamation points the same way.

当 \TeX\ 看见一个句号后面紧接着一个空格（或者是一个行结束符，这
两者等价）时，它通常会认为此句号代表一句话的结束，并在句号的后面
加上一点小小的空格。同样的处理还适用于问号和惊叹号。

%  But \TeX's rules for handling periods sometimes need fine
%tuning. \TeX\ assumes that a capital letter before a period
%doesn't end the sentence, so you have to do something a little
%different if, say, you're writing about DNA\null.

有时需要微调 \TeX 处理句号的方式。当句号前面紧跟着标题字
母（大写字母）时，\TeX\ 会认为这个句号并不做为句子的结束符，
因此在这种情况下，你需要作一些小改动。例如：我们这样写 DNA\null.


% The \null prevents TeX from perceiving the capital `A'
% as being next to the period.
% \null 可以阻止 \TeX\ 感知句号前面的标题字母 `A'。

%It's a good idea to tie words together in references such as
%``see Fig.~8'' and in names such as V.~I\null. Lenin and in
%$\ldots$ so that \TeX\ will never split them in an awkward place
%between two lines.  (The three dots indicate an ellipsis.)

% 这个 references 不知道这样翻译对不对
如果想将多个单词绑定到一起，中间不分行，则可以将它们放入一个引用
中，例如： ``see Fig.~8''；或者名字中，例如：V.~I\null. Lenin；或者
省略号 $\ldots$。在这样的情况下，\TeX\ 不会在它们之间断行（三个点表示
省略号）。

You should put quotations in pairs of left and right
single ``quotes'' so that you get the correct left and right
double quotation marks.  ``For adjacent single and double
quotation marks, insert a `thinspace'\thinspace''. You can
get en-dashes--like so, and em-dashes---like so.

\bye % end the document
:::
\xmpheader !xmpnum/{Indentation}% !xmpheaddef
~^^{indentation} ^^{margins} ^^{paragraphs//narrow}
%\noindent Now let's see how to control indentation.  If an
%ordinary word processor can do it, so surely can \TeX. Note
%that this paragraph isn't indented.

\noindent 现在让我们来看看如何控制缩进。如果一个普通的文字处理
程序都能处理缩进的话，那 \TeX\ 一定也能。请注意，本段并未缩进。

%Usually you'll either want to indent paragraphs or to leave
%extra space between them.  Since we haven't changed anything
%yet, this paragraph is indented.

通常来说，你会想在段落的起始处安排缩进，在段落之间安排一些额外
的空白。由于我们在此尚未做任何特殊处理，因此本段是缩进的。

{\parindent = 0pt \parskip = 6pt
%% The left brace starts a group containing the unindented text.
% 这里的左括弧包裹了一组不缩进的文本

%Let's do these two paragraphs a different way,
%with no indentation and six printer's points of extra space
%between paragraphs. 

让我们对这两个段落作一些特殊处理，这两段将不缩进，并且两段
之间有 6pt 宽的空白。

%So here's another paragraph that we're typesetting without
%indentation.  If we didn't put space between these paragraphs,
%you would have a hard time knowing where one ends
%and the next begins.

这是第二段不缩进的文字。如果不在这两段之间插入空白的话，将很难
分辨段与段之间的结束和起始。

\par % The paragraph *must* be ended within the group.
}% The right brace ends the group containing unindented text.

%It's also possible to indent both sides of entire paragraphs.  
%The next three paragraphs illustrate this:

你也能在段落的两边同时缩进，下面的 3 段文字演示了双边缩进的效果：


\smallskip % Provide a little extra space here.
% Skips like this and \vskip below end a paragraph.
{\narrower
% ``We've indented this paragraph on both sides by the paragraph
% indentation.  This is often a good way to set long quotations.''

``我们对这一段作了双边缩进的处理。在长段引用时，通常都会这么做。''

%``You can do multiple paragraphs this way if you choose.  This
%is the second paragraph that's singly indented.''\par}

``如果你愿意的话，可以同时对多个段落作如此处理.  这是
第二个轻微缩进的段落.''\par}

%{\narrower \narrower You can even make paragraphs doubly narrow
%if that's what you need.  This is an example of a doubly 
%narrowed paragraph.\par}
{\narrower \narrower 如果需要，你也可以让段落缩进双倍长度。
这就是一个双倍缩进段落的例子.\par}
\vskip 1pc % Skip down one pica for visual separation.
In this paragraph we're back to the normal margins, as you can
see for yourself.  We'll let it run on a little longer so that
the margins are clearly visible.

在这一段，如你所见，我们返回到正常的边距状态。我们尽量让这一段
稍微长一点，这样你可以更清楚地看到边距的效果。

%{\leftskip .5in Now we'll indent the left margin by half
%an inch and leave the right margin at its usual position.\par}

{\leftskip .5in 现在这一段，我们将其调整为左边距为半英尺，而
右边距保持原样。\par}


%{\rightskip .5in Finally, we'll indent the right margin by half
%an inch and leave the left margin at its usual position.\par}
{\rightskip .5in 最后，我们将本段调整为右边距半英尺，而左边距保持
原样。\par}
\bye % end the document
:::
\xmpheader !xmpnum/{Fonts and special characters}% !xmpheaddef
\chardef \\ = `\\ % Let \\ denote a backslash.
~^^{fonts} ^^{characters//special} ^^{accents}
~^^{music symbols} ^^{card suits}
~^^|$| ^^|&| ^^|#| ^^|_| ^^|%| ^^|^| ^^|~| ^^|{| ^^|}| \indexchar \
%{\it Here are a few words in an italic font}, {\bf a 
%few words in a boldface font}, {\it and a\/ {\bf mixture} 
%of the two, with two\/ {\rm roman words} inserted}.
%Where an italic font is followed by a nonitalic font, we've
%inserted an ``italic correction'' ({\tt \\/}) to make the
%spacing look right. 

%{\it 这是一些斜体字词}, {\bf  
%一些粗体字词}, {\it 以及一些 {\bf 二者混合的效果} 
%, 我们在中间插入一些 {\rm 正常的字词} }.
%当斜体字后面紧跟非斜体字时, 我们在其间插入一个``倾斜校正''
%({\tt \\/})，使得中间的空白看起来较舒适。

%Here's a {\sevenrm smaller} word---but the standard \TeX\
%fonts won't give you anything smaller than {\fiverm this}.

这是一些{\sevenrm 更小的} 字词---可是标准的 \TeX\
字体并不会给出小于 {\fiverm 如此大小}的字体。

%If you need any of the ten characters:
如果你需要以下十个字符任意之一：
\medskip
\centerline{\$ \quad \& \quad \# \quad \_ \quad \% \quad
   \char `\^ \quad \char `\~ \quad $\{$ \quad
   $\}$ \quad $\backslash$}
% The \quad inserts an em space between characters.
\medskip
%\noindent  you'll need to write them a special way.  Look at
%the facing page to see how to do it.
\noindent 你需要通过特殊的方式来写他们。参照首页，学习正确的记录
这些特殊字符的方法.

%\TeX\ has the accents and letters that you'll need
%for French words such as {\it r\^ ole\/} and  {\it \'
%el\` eve\/}, for German words such as {\it Schu\ss\/},
%and for words in several other languages as well.
%You'll find a complete list of \TeX's accents and letters
%of European languages on !xrefdelim[accents] and !xrefdelim[fornlets].

\TeX\ 拥有一些语音和字母，可以用于某些法语词，例如：{\it r\^ ole\/} 以及
{\it \'el\` eve\/}, 或者德语词，例如 {\it Schu\ss\/},
或者某些别的什么语言。你可以在 !xrefdelim[accents] 和 !xrefdelim[fornlets] 处
找到 \TeX\ 的 欧洲语音和字母表。

%You can also get Greek letters such as ``$\alpha$'' and
%``$\Omega$'' for use in math, card suits such as
%``$\spadesuit$'' and ``$\diamondsuit$'', music symbols
%such as ``$\sharp$'' and ``$\flat$'', and many other special
%symbols that you'll find listed on !xrefdelim[specsyms].
%\TeX\ will only accept these sorts of special symbols in its
%``math mode'', so you'll need to enclose them
%within `{\tt \$}' characters.

你也可以写一些希腊字母，例如数学中的 ``$\alpha$'' 以及
``$\Omega$'' , 扑克牌中的 
``$\spadesuit$'' 以及 ``$\diamondsuit$'', 一些音乐符号，例如：
``$\sharp$'' 和 ``$\flat$'', 以及一些你可以在 !xrefdelim[specsyms]
找到的各种符号。\TeX\ 仅在 ``数学模式'' 下接受这些符号, 因此，
你需要将它们放在两个 `{\tt \$}' 字符中间。
\bye % end the document
:::
\xmpheader !xmpnum/{Interline spacing}% !xmpheaddef
~^^{spacing//interline} ^^{baselines}
\baselineskip = 2\baselineskip % double spacing
\parskip = \baselineskip % Skip a line between paragraphs.
\parindent = 3em % Increase indentation of paragraphs.

% The following macro definition gives us nice inline
% fractions.  You'll find it in our eplain macros.
\def\frac#1/#2{\leavevmode
   \kern.1em \raise .5ex \hbox{\the\scriptfont0 #1}%
   \kern-.1em $/$%
   \kern-.15em \lower .25ex \hbox{\the\scriptfont0 #2}%
}%

%Once in a while you may want to print a document with extra
%space between the lines.  For instance, bills before Congress
%are printed this way so that the legislators can mark them up.
%For the same reason, book publishers usually insist that
%manuscripts be double-spaced. Double spacing is rarely
%appropriate for finished documents, however.

有的时候，你会想在文档的文字行之间增加一些空白。例如：议会的
某些讨论稿需要这样做，使得议院们能够在其上作一些标记。书商们
也基于同样的理由，要求作者提交的手稿具有双倍的行间距。当然，
对最终的出版物来说，双倍行间距的情形极其少见。


%A baseline is an imaginary line that acts like the lines
%on a pad of ruled paper.   You can control the interline
%spacing---what printers call ``leading''---%
%by setting the amount of space between baselines. Take a
%look at the input to see how to do it. You could use
%the same method for $1\;1/2$ spacing as well, using {\tt 1.5}
%instead of {\tt 2}. (You can also write $1\frac 1/2$
%a nicer way.)

基线是一条想象出来的线，其作用类似于带横格标记的纸张上
横格线的作用。你可以通过控制基线间的距离来控制两行文字
之间的距离 --- 打印机管这个距离叫做 ``起始空白''。具体
的做法可以参看这段的源文件。你可以用{\tt 1.5}取代{\tt 2}
获得 $1\;1/2$空白的效果。（或者可以写成更好看的 $1\frac 1/2$
形式。）

% Here we've used the macro definition given above.

%For this example we've also increased the paragraph indentation
%and skipped an extra line between paragraphs.

做为举例，我们这里增加了段落的缩进，并且在两段之间增加了一个空行。

\bye % end the document
:::
\xmpheader !xmpnum/{Spacing, rules, and boxes}% !xmpheaddef
~^^{description lists} ^^{boxes//drawing} ^^{revision bars}
%Here's an example of a ``description list''.  In practice you'd 
%be better off using a macro to avoid the repetitive constructs
%and to make sure that the subhead widths are wide enough:

这里展示的是 ``描述列表''的实例。实际使用时，最好将这些重复结构
定义成宏，并且保证子标题的宽度足够宽。

\bigskip
% Call the indentation for descriptions \descindent
% and set it to 8 picas.
\newdimen\descindent \descindent = 8pc
% Indent paragraphs by \descindent.
% Skip an additional half line between paragraphs.
{\noindent \leftskip = \descindent  \parskip = .5\baselineskip
% Move the description to the left of the paragraph.
\llap{\hbox to \descindent{\bf Queen of Hearts\hfil}}%
An ill-tempered woman, prone to saying ``Off with his
head!!''\ at the slightest provocation.\par
\noindent\llap{\hbox to \descindent{\bf Cheshire Cat\hfil}}%
A cat with an enormous smile that Alice found
in a tree.\par
\noindent\llap{\hbox to \descindent{\bf Mock Turtle\hfil}}%
A lachrymose creature, quite a  storyteller, who was a
companion to the Gryphon.  Reputedly the principal ingredient
of Mock Turtle Soup.
\par}
\bigskip\hrule\bigskip % A line with vertical space around it.
Here's an example of some words in a ruled box, just as
Lewis Carroll wrote them:
\bigskip
% Put 8pt of space between the text and the surrounding rules.
\hbox{\vrule\vbox{\hrule
   \hbox spread 8pt{\hfil\vbox spread 8pt{\vfil
      \hbox{Who would not give all else for twop}%
      \hbox{ennyworth only of Beautiful Soup?}%
   \vfil}\hfil}
\hrule}\vrule}%

\bigskip\line{\hfil\hbox to 3in{\leaders\hbox{ * }\hfil}\hfil}
\bigskip

\line{\hskip -4pt\vrule\hfil\vbox{
Here we've gotten the effect of a revision bar on the material
in this paragraph.  The revision bar might indicate a change.}}
\bye % end the document
:::
\xmpheader !xmpnum/{Odds and ends}% !xmpheaddef
~^^{hyphenation} ^^{theorems} ^^{lemmas} ^^{itemized lists}
~^^{flush left} ^^{flush right} ^^{centering}
\chardef \\ = `\\ % Let \\ denote a backslash.
\footline{\hfil{\tenit - \folio -}\hfil}
~\global\footline{\hfil{\tenit - \folio\ -}\hfil}
% \footline provides a footer line.
% Here it's a centered, italicized page number.
\TeX\ knows how to hyphenate words, but  it isn't infallible.
If you are discussing the chemical
${\it 5}$-[p-(Flouro\-sul\-fonyl)ben\-zoyl]-l,%
$N^6$-ethe\-no\-adeno\-sine
and \TeX\ complains to you about an ``overfull hbox'', try
inserting some ``discretionary hyphens''. The notation
`{\tt \\-}' tells \TeX\ about a  dis\-cre\-tion\-ary hyphen,
that is, one that it might not have inserted otherwise.
\medskip
{\raggedright   You can typeset text unjustified, i.e., with
an uneven right margin. In the old days, before word
processors were common, typewritten documents were
unjustified because there was no convenient alternative.
Some people  prefer text to be  unjustified so that the
spacing between words can be uniform.  Most books are set
with justified margins, but not all. \par}

\proclaim Assertion 27. There is an easy way to typeset
the headings of assertions, lemmas, theorems, etc.

Here's an example of how to typeset an itemized list two
levels deep.  If you need more levels, you'll have to
program it yourself, alas.
\smallskip
\item {1.} This is the first item.
\item {2.} This is the second item.  It consists of two
paragraphs.  We've indented the second paragraph so that
you can easily see where it starts.

\item{} \indent The second paragraph has three subitems
underneath it.
\itemitem {(a)} This is the first subitem.
\itemitem {(b)} This is the second subitem.
\itemitem {(c)} This is the third subitem.
\item {$\bullet$} This is a strange-looking item because it's
completely different from the others.
\smallskip
\leftline{Here's a left-justified line.$\Leftarrow$}
\rightline{$\Rightarrow$Here's a right-justified line.}
\centerline{$\Rightarrow$Here's a centered line.$\Leftarrow$}
% Don't try to use these commands within a paragraph.
\bye % end the document
:::
\xmpheader !xmpnum/{Using fonts from other sources}% !xmpheaddef
~\xrdef{palatino}
~\idxref{Palatino fonts}
~\idxref{Zapf, Hermann}
~\idxref{Computer Modern fonts} ^^{\Metafont}
\font\tenrm = pplr % Palatino
% Define a macro for invoking Palatino.
\def\pal{\let\rm = \tenrm \baselineskip=12.5pt \rm}
\pal % Use Palatino from now on.

%You aren't restricted to using the Computer Modern fonts that
%come with \TeX.  Other fonts are available from many sources,
%and you may prefer them.  For instance, we've set this page
%in 10-point Palatino Roman. Palatino was designed by
%Hermann Zapf, considered to be one of the greatest type
%designers of the twentieth century.  This page will
%give you some idea of what it looks like.

你并不局限于只能使用随 \TeX 发行的 Computer Modern 系列字体。
从其他很多渠道你都可以得到一些你可能会更喜欢的字体。例如我们
当前页面就被设置为 10p 大小的 Palatino Roman 字体。Palatino
是一种由 Hermann Zapf 所设计的字体，它被认为是二十世纪最伟大的
字体设计之一。从本页的输出可以获得关于这个字体的一些直观印
象（由于译本的缘故，本页的字体并未被设为 Palatino 字体）。

%Fonts can be provided either as outlines or as bitmaps.  An
%outline font describes the shapes of the characters, while a
%bitmap font specifies each pixel (dot) that makes up each
%character.  A font outline can be used to generate many
%different sizes of the same font.  The Metafont program
%that's associated with \TeX\  provides a particularly
%powerful way of generating bitmap fonts, but it's not the
%only way.

字体既可以是轮廓字体，也可以是位图字体。轮廓字体的意思是它
描述的是各个字符的形状，而位图字体则是标识了每个字符形状所
占据的像素点。轮廓字体可用来产生同一种字体的多种不同大小。
随 \TeX\ 而发行的 metafont 程序是一个产生位图字体的强大工具，
而且它并不是唯一的工具。

%The fact that a single outline can generate a great range of
%point sizes for a font tempts many vendors of digital
%typefaces to provide just one set of outlines for a typeface
%such as Palatino Roman.  This may be a sensible economic
%decision, but it is an aesthetic sacrifice.  Fonts cannot be
%scaled up and down linearly without loss of quality. 
%Larger sizes of letters should not, in general, have the
%same proportions as smaller sizes; they just don't look
%right.  For example, a font that's linearly scaled down will
%tend to have too little space between strokes, and its
%x-height will be too~small. % tie added to avoid widow word

由于单一的轮廓字体文件既可生成多种不同大小的字体，因此很多
数字字体提供商倾向于仅提供字体的单一文件 ——例如 Palatino 
Roman 即是如此。这也许是个经济的决定，可它却同时是个美学上
的重大牺牲。字体无法做到在不损害其显示质量的前提下，线性地
放大或者缩小。一个字符的大字体版本，通常来说，与小字体的版
本应该有不同的比例 —— 而如果我们坚持这么做的话，其结果看
起来将不会那么好看。例如：如果将一个字体线性缩小的话，其各
笔画间的间距将会显得过窄，而它的 x-height 值也会显得过小。

%A type designer can compensate for these changes by
%providing different outlines for different point sizes, but
%it's necessary to go to the expense of designing these
%different outlines.  One of the great advantages of Metafont
%is that it's possible to parameterize the descriptions of
%characters in a font.  Metafont can then maintain the
%typographical quality of characters over a range of point
%sizes by adjusting the character shapes accordingly.

字体的设计者可以为不同大小的字体设计不同的轮廓形状，从而
对这个问题进行弥补，可这同时需要付出额外的代价。使用 Metafont
程序的一个巨大的好处就是可以将字符符号形状的描述参数化。
这样在改变字体大小的时候，Metafont 可以视情况做一些自动处理，
从而得到较好的效果。
\bye % end the document
:::
~\idxref{mushrooms}
\xmpheader !xmpnum/{A ruled table}% !xmpheaddef
\bigskip
\offinterlineskip % So the vertical rules are connected.
% \tablerule constructs a thin rule across the table.
\def\tablerule{\noalign{\hrule}}
% \tableskip creates 9pt of space between entries.
\def\tableskip{\omit&height 9pt&&&\omit\cr}
% & separates templates for each column. TeX substitutes
% the text of the entries for #. We must have a strut
% present in every row of the table; otherwise, the boxes
% won't butt together properly, and the rules won't join.
\halign{\tabskip = .7em plus 1em  % glue between columns
% Use \vtop for short multiline entries in the first column.
% Typeset the lines ragged right, without hyphenation.
   \vtop{\hsize=6pc\pretolerance = 10000\hbadness = 10000
      \normalbaselines\noindent\it#\strut}%
  &\vrule #&#\hfil &\vrule #% the rules and middle column
% Use \vtop to get whole paragraphs in the last column.
  &\vtop{\hsize=11pc \parindent=0pt \normalbaselineskip=12pt
    \normalbaselines \rightskip=3pt plus2em #}\cr
% The table rows begin here.
\noalign{\hrule height2pt depth2pt \vskip3pt}
  % The header row spans all the columns.
  \multispan5\bf Some Choice Edible Mushrooms\hfil\strut\cr
\noalign{\vskip3pt} \tablerule
  \omit&height 3pt&\omit&&\omit\cr
  \bf Botanical&&\bf Common&&\omit \bf Identifying \hfil\cr
\noalign{\vskip -2pt}% close up lines of heading
  \bf Name&&\bf Name &&\omit \bf Characteristics \hfil\cr
\tableskip Pleurotus ostreatus&&Oyster mushroom&&
  Grows in shelf\kern 1pt like clusters on stumps or logs,
  % without the kern, the `f' and `l' would be too close
  pink-gray oyster-shaped caps, stem short or absent.\cr
\tableskip Lactarius hygrophoroides&&Milky hygroph&&
  Butterscotch-brown cap and stem, copious white latex,
  often on ground in woods near streams.\cr
\tableskip Morchella esculenta&&White morel&&Conical cap
  with black pits and white ridges; no gills. Often found
  near old apple trees and dying elms in the spring.\cr
\tableskip Boletus edulus&&King bolete&&Reddish-brown to
  tan cap with yellow pores (white when young),
  bulbous stem, often near conifers, birch, or~aspen.\cr
\tableskip \tablerule \noalign{\vskip 2pt} \tablerule
}\bye
:::
\xmpheader !xmpnum/{Typesetting mathematics}% !xmpheaddef
~^^{math}
For a spherical triangle with sides $a$, $b$, and $c$, and
opposite angles $\alpha$, $\beta$, and $\gamma$, we have:
$$\cos \alpha = -\cos \beta \cos \gamma +
  \sin \beta \sin \gamma \cos \alpha \quad
  \hbox{(Law of Cosines)}$$
and:
$$\tan {\alpha \over 2} = \sqrt{
  {- \cos \sigma \cdot \cos(\sigma - \alpha)} \over
  {\cos (\sigma - \beta) \cdot \cos (\sigma - \gamma)}},\quad
  \hbox{where $\sigma = {1 \over 2}(a+b+c)$}$$
We also have:$$\sin x = {{e^{ix}-e^{-ix}}\over 2i}$$
and:
$$\int _0 ^\infty {{\sin ax \sin bx}\over{x^2}}\,dx
% The \, above produces a thin space
  = {\pi a\over 2}, \quad \hbox{if $a < b$}$$

\noindent The number of combinations ${}_nC_r$ of $n$
things taken $r$ at a time is:
$$C(n,r) = {}_nC_r = {n \choose r} =
  {{n(n-1) \cdots (n-r+1)} \over {r(r-1) \cdots (1)}} =
  {{n!!}\over {r!!(n-r)!!}}$$

\noindent
The value of the determinant $D$ of order $n$:
$$D = \left|\matrix{a_{11}&a_{12}&\ldots&a_{1n}\cr
  a_{21}&a_{22}&\ldots&a_{2n}\cr
  \vdots&\vdots&\ddots&\vdots\cr
  a_{n1}&a_{n2}&\ldots&a_{nn}\cr}\right| $$
is defined as the sum of $n!!$ terms:
$$\sum\>(\pm)\>a_{1i}a_{2j} \ldots a_{nk}$$
% The \> above produces a medium space.
where $i$, $j$, \dots,~$k$\/ take on all possible values
between $1$ and $n$, and the sign of the product is 
$+$ if the sequence $i$, $j$, \dots,~$k$\/ is an
even permutation and $-$ otherwise.  Moreover:
$$Q(\xi) = \lambda_1 y_1^2 \sum_{i=2}^n \sum_{j=2}^n y_i
b_{ij} y_j,\qquad B = \Vert b_{ij} \Vert = B'$$
\bye
:::
%\xmpheader !xmpnum/{More mathematics}% !xmpheaddef
\xmpheader !xmpnum/{更多数学内容}% !xmpheaddef
%^^{math}
^^{数学}
%The absolute value of $X$, $|x|$, is defined by:
%$$|x| = \cases{x, &if $x\ge 0$;\cr
%-x,&otherwise.\cr}$$
%Now for some numbered equations.
%It is the case that for $k \ge 0$:
%$$x^{k^2}=\overbrace{x\>x\>\cdots\> x}^{2k\ \rm times}
%\eqno (1)$$

$X$ 的绝对值 $|x|$ 定义为：:
$$|x| = \cases{x, &if $x\ge 0$;\cr
-x,&otherwise.\cr}$$
现在写一些标号了的公式。
在 $k \ge 0$ 时，我们有：
$$x^{k^2}=\overbrace{x\>x\>\cdots\> x}^{2k\ \rm times}
\eqno (1)$$

%Here's an example that shows some spacing controls, with
%a number on the left:
这是一个展示空格间距调整和公式左编号的例子：
%$$[u]!negthin[v][w]\,[x]\>[y]\;[z]\leqno(2a)$$
$$[u]!negthin[v][w]\,[x]\>[y]\;[z]\leqno(2a)$$
%The amount of space between the items in brackets
%gradually increases from left to right. (We've made
%the space between the first two items be {\it less\/}
%than the natural space.)
从做到右公式中的项和括号的距离逐步增大。
（我们把前两项的距离设定得比正常距离 {\it 少\/}。
%It is sometimes the case that $$\leqalignno{
%u'_1 + tu''_2 &= u'_2 + tu''_1&(2b)\cr
%\hat\imath &\ne \hat \jmath&(2c)\cr
%\vec {\vphantom{b}a}&\approx \vec b\cr}$$
%% The \vphantom is an invisible rule as tall as a `b'.
有时候， $$\leqalignno{
u'_1 + tu''_2 &= u'_2 + tu''_1&(2b)\cr
\hat\imath &\ne \hat \jmath&(2c)\cr
\vec {\vphantom{b}a}&\approx \vec b\cr}$$
% \vphantom 是一个不可见的和“b”一样高的盒子。
%The result is of order $O(n \log\log n)$. Thus
%$$\sum_{i=1}^n x_i = x_1+x_2+\cdots+x_n
%= {\rm Sum}(x_1,x_2,\ldots,x_n). \eqno(3)$$
结果为 $O(n \log\log n)$. 因此
$$\sum_{i=1}^n x_i = x_1+x_2+\cdots+x_n
= {\rm Sum}(x_1,x_2,\ldots,x_n). \eqno(3)$$
%and
%$$dx\,dy = r\,dr\,d\theta!negthin.\eqno(4)$$
%The set of all $q$ such that $q\le0$ is written as:
%$$\{\,q\mid q\le0\, \}$$
且
$$dx\,dy = r\,dr\,d\theta!negthin.\eqno(4)$$
满足 $q\le0$ 的所有的 $q$ 可以被写作：
$$\{\,q\mid q\le0\, \}$$
%Thus
%$$\forall x\exists y\;P(x,y)\Rightarrow
%\exists x\exists y\;P(x,y)$$
%where
%$$P(x,y) \buildrel \rm def \over \equiv
%\hbox{\rm any predicate in $x$ and $y$} . $$
因此
$$\forall x\exists y\;P(x,y)\Rightarrow
\exists x\exists y\;P(x,y)$$
其中
$$P(x,y) \buildrel \rm def \over \equiv
\hbox{\rm 任何一个能够得出 $x$ 和 $y$} . $$
\bye
:::
